---
title: "2 - Paper Figures"
output:
  html_document:
    df_print: paged
---
Final Figures for CMEJ Paper.  

# Setup

## Library

```{r}
library(extrafont)

## import fontawesome. If you have to redo, download ttf directly from emojifont package
# font_import(paths="C:/Users/azeng/AppData/Local/Microsoft/Windows/Fonts/", pattern = "fontawesome-webfont")
# font_import(paths="C:/Windows/Fonts/", pattern = "fontawesome-webfont")

# check that Font Awesome is imported
fonts()[grep("Awesome", fonts())]
# [1] "FontAwesome"

# use this if things look odd in RStudio under Windows
loadfonts(device = "win")
```

Libraries
```{r, warning=FALSE}
library(ggpubr)
library(ggbiplot)
library(dendextend)
library(Rtsne)
library(ggthemes)
library(tidyverse)
library(waffle)
library(RColorBrewer)
library(factoextra)
```


## Data

Load Electives with cluster assignments 
```{r}
carms <- read.csv("carms_cleaned.csv")
electives <- read.csv("electives.csv", stringsAsFactors = FALSE)

electives_clust <- tibble("A" = c("Microbio", "Gen Path", "Nuclear", "Pub Health","Neuro Path","Peds Neuro",
                          "Heme Path", "Anat Path","Internal","Family", "Med Gen", NA),
                  "B" = c("Psych", "Rad Onc", "OBGYN", "Peds", "Neuro", "Physiatry", "Anesth",
                          "Rads", "Gen Surg", "Vasc Surg", "Derm", "Emerg"),
                  "C" = c("Cardiac Surg", "Ophthal",  "Neuro Surg", "Urology",
                          "Ortho", "ENT", "Plastics", NA, NA, NA, NA, NA)) %>%
 gather(key = "cluster", value = "disc_code") %>% na.omit() %>%
 right_join(.,electives, by="disc_code")


# approximate simpsons diversity index for electives diversity metric
  # inflate everything by ten to get clean integer
get_simp <- function(dat){
  num_other_disciplines <- 10*dat$mean_other_disciplines
  simp <- (dat$mean_electives_within_discipline / (dat$mean_electives_within_discipline + dat$mean_electives_outside_discipline))^2
  
  for(i in 1:num_other_disciplines){
    simp <- simp + ((dat$mean_electives_outside_discipline / dat$mean_other_disciplines) / (10*(dat$mean_electives_within_discipline + dat$mean_electives_outside_discipline)))^2
  }
  dat$electives_diversity <- 1 - simp
  return(dat)
}

# get electives diversity
electives_clust <- electives_clust %>% 
  filter(year >= 2013) %>%
  na.omit() %>%
  mutate(discipline_year = paste0(disc_code, " (", year, ")")) %>%
  split(.$discipline_year) %>% 
  map_dfr(~ get_simp(.))

electives_clust <- carms %>% 
  mutate(disc_code = as.character(disc_code)) %>%
  mutate(disc_code = if_else(disc_code == "Opthal", "Ophthal", disc_code)) %>%
  mutate(p_nomatch_unmatch = unmatched / (matched_to_other + unmatched),
         p_nomatch_other = 1 - p_nomatch_unmatch, 
         app_as_first_choice = app_choice_1 / app_total)%>% 
  select(year, disc_code, p_nomatch_unmatch, p_nomatch_other, app_as_first_choice, app_choice_1, seat_total, seat_vacant, matched_choice_1, matched_to_other, unmatched) %>% 
  left_join(electives_clust, ., by = c("year", "disc_code"))


electives_clust %>% head()

```

# Main Figures

## Figure 1

Ratios in first iteration and introduction to the clusters


### Ratio plot for first iteration


```{r, include=FALSE}
ratio_palette <- c("#BDBDBD", "dodgerblue4", "firebrick4")

discipline_trend <- carms

# discipline_trend with cluster labels
discipline_trend_clust <- electives_clust %>%
  mutate(cluster = paste0("Cluster ", cluster)) %>%
  dplyr::select(discipline_full, cluster) %>%
  unique() %>% 
  left_join(discipline_trend, ., by = c("discipline_full"))
```


```{r}
app_seat_iter_1 <- discipline_trend_clust %>% 
  group_by(year) %>% 
  dplyr::summarise(`CMG Applicants` = sum(app_choice_1), 
                   `Residency Seats` = sum(seat_total),
                   `Unmatched` = sum(unmatched)) %>% 
  ungroup() %>%
  gather(key = "First", value = "Count", `CMG Applicants`, `Residency Seats`, `Unmatched`) %>%
  dplyr::select(year, First, Count)

app_seat_1 <- ggplot() +
  geom_col(data = filter(app_seat_iter_1, First == "Residency Seats"), 
           aes(x = year, y = Count, fill = " Residency Seats ")) +
  geom_line(data = filter(app_seat_iter_1, First == "CMG Applicants"), 
            aes(x = year, y = Count, color = " CMG Applicants     "), size = 2) +
  geom_line(data = filter(app_seat_iter_1, First == "Unmatched"), 
            aes(x = year, y = Count, color = " Unmatched"), size = 1.5) +
  ylab("Count") + xlab("Year") +
  scale_x_continuous(breaks=seq(2010, 2019,2)) +
  scale_fill_manual(name = "", values = c(" Residency Seats " = ratio_palette[1])) +
  scale_color_manual(name = "", values = c(" CMG Applicants     " = ratio_palette[2],
                                           " Unmatched" = ratio_palette[3])) +
  theme_pubr() +
  theme(legend.title = element_blank(),
        legend.position = "bottom") 


first_iter <- data.frame("First Iteration Match - All Disciplines")

ggarrange(ggtexttable(first_iter, rows=NULL, cols=NULL, 
                      theme = ttheme("lBlack", padding = unit(c(100,5),"mm"), base_size = 14)),
          app_seat_1 + ylab("Count") + theme(plot.margin = unit(c(1,0.5,1.5,0.5), "lines"), legend.position = 'none'), 
          heights = c(0.3, 2.02), ncol = 1, nrow = 2)
```

## Load summarised discipline data (median from 2013-2019)

```{r}
cluster_palette <- c("#46B8DACC", "#EEA236BB", "#9632B8BB")
tsne_data <- read_csv("tsne_electives.csv")

tsne_data %>% head()
```

## Cluster Dendrogram

```{r}
electives_mat <- tsne_data %>% select(-c("discipline", "discipline_full", "disc_code", "prop_unmatched", "TSNE1", "TSNE2")) %>% data.matrix()
rownames(electives_mat) <- tsne_data %>% pull(disc_code)

# get dendrogram
corelectives <- electives_mat %>% t() %>% cor()
electives_dist <- as.dist((1-corelectives)/2)
electives_hc <- hclust(electives_dist, method = "complete")

dend = fviz_dend(rev(electives_hc), k=3, labels_track_height = 0.05, main = "", 
                 ylab="Distance", k_colors=cluster_palette)
dend
```


### Cluster TSNE plot


```{r}
tsne_plot <- tsne_data %>% 
    mutate(Cluster = if_else(electives_clusters == 1, "B", if_else(electives_clusters == 2, "A", "C"))) %>%
    ggplot(aes(TSNE1, TSNE2, color = Cluster)) +
    geom_text(aes(label = disc_code), size = 3.2) + theme_pubr() +
    theme(legend.position = "top", plot.margin = unit(c(1, 1,2,1), "lines"), 
          axis.text = element_blank(), axis.ticks = element_blank(), 
          legend.title=element_text(size=14), 
          legend.text = element_text(size=12)) +
    scale_color_manual(values = cluster_palette) +
    guides(color = guide_legend(override.aes = list(size=5.4)))

tsne_plot
```

### Ratio plot by cluster

```{r}
app_seat_disc_class <- discipline_trend_clust %>% 
  group_by(year, cluster) %>% 
  dplyr::summarise(`CMGs Ranking Discipline as First Choice` = sum(app_choice_1), 
                   `Residency Seats` = sum(seat_total),
                   `Unmatched Applicants` = sum(unmatched)) %>% 
  ungroup() %>%
  gather(key = "First", value = "Count", `CMGs Ranking Discipline as First Choice`, `Residency Seats`, `Unmatched Applicants`) %>%
  dplyr::select(year, cluster, First, Count)
```

```{r, fig.width=14, fig.height=5}
ratio_palette <- c("#BDBDBD", "dodgerblue4", "firebrick4")

plot_ratio <- function(app_seat_disc_class, cluster_name, palette){
  spots_clust <- ggplot() +
    geom_col(data = filter(app_seat_disc_class, First == "Residency Seats") %>% filter(cluster == cluster_name),
             aes(x = year, y = Count, fill = " Residency Seats   ")) +
    geom_line(data = filter(app_seat_disc_class, First == "CMGs Ranking Discipline as First Choice") %>% filter(cluster == cluster_name),
             aes(x = year, y = Count, color = " CMGs Ranking Discipline as First Choice    "), size = 1.8) +
    geom_line(data = filter(app_seat_disc_class, First == "Unmatched Applicants") %>% filter(cluster == cluster_name), 
             aes(x = year, y = Count, color = " Unmatched Applicants"), size = 1.8) +
    scale_x_continuous(breaks=seq(2010, 2018,2)) +
    scale_fill_manual(name = "", values = c(" Residency Seats   " = palette[1])) +
    scale_color_manual(name = "", values = c(" CMGs Ranking Discipline as First Choice    " = palette[2],
                                             " Unmatched Applicants" = palette[3])) +
    theme_pubr() + theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
    facet_wrap(~cluster, scales = "free_y")
  
  return(spots_clust)
}

# cluster A
spots_clust_A <- plot_ratio(app_seat_disc_class, "Cluster A", ratio_palette)
# cluster B
spots_clust_B <- plot_ratio(app_seat_disc_class, "Cluster B", ratio_palette)
# cluster C
spots_clust_C <- plot_ratio(app_seat_disc_class, "Cluster C", ratio_palette)


ggarrange(
  spots_clust_A + theme(strip.text.x = element_text(size = 13), plot.margin = unit(c(1,0.5,0,0.5), "lines")), 
  spots_clust_B + theme(strip.text.x = element_text(size = 13), plot.margin = unit(c(1,0.5,0,0.5), "lines")),
  spots_clust_C + theme(strip.text.x = element_text(size = 13), plot.margin = unit(c(1,0.5,0,0.5), "lines")),
  nrow = 1, ncol = 3, labels = c("C", "D", "E"), font.label = list(size = 22),
  common.legend = TRUE, legend = "bottom"
)
```


```{r, fig.width=14, fig.height=9}
first_iter <- data.frame("First Iteration Match - All Disciplines")

ggarrange(
  ggarrange(dend, tsne_plot,
          ncol = 2, nrow = 1, widths = c(1, 1.3), labels = c("A", "B"), font.label = list(size = 24)),
  ggarrange(
    spots_clust_A + theme(strip.text.x = element_text(size = 13), plot.margin = unit(c(1,0.5,0,0.5), "lines")), 
    spots_clust_B + theme(strip.text.x = element_text(size = 13), plot.margin = unit(c(1,0.5,0,0.5), "lines")),
    spots_clust_C + theme(strip.text.x = element_text(size = 13), plot.margin = unit(c(1,0.5,0,0.5), "lines")),
    nrow = 1, ncol = 3, labels = c("C", "D", "E"), font.label = list(size = 22),
    common.legend = TRUE, legend = "bottom"
  ), nrow = 2, ncol = 1, heights = c(1.3,1)) + ggsave("paper_figures/Figure1_Clusters_and_ratios.png", height = 9, width = 14, dpi = 600)
```




###############################################################################################################################################

## Figure 2

Match Stats of disciplines (no electives) and match outcomes barplot 

```{r}
electives_clust <- electives_clust %>% filter(cluster != "NA")
```

```{r}
cluster_labels <- c("A" = "Cluster A", "B" = "Cluster B", "C" = "Cluster C")

outcomes <- electives_clust %>% 
  filter(year >= 2013) %>% 
  group_by(disc_code, cluster) %>% 
  dplyr::summarise(app_choice_1 = sum(app_choice_1), 
            matched_choice_1 = sum(matched_choice_1),
            matched_to_other = sum(matched_to_other),
            unmatched = sum(unmatched)) %>% 
  mutate(prop_matched_to_discipline = matched_choice_1 / app_choice_1,
         prop_matched_to_alternative = matched_to_other / app_choice_1,
         prop_unmatched = unmatched / app_choice_1) %>% 
  select(disc_code, cluster, contains("prop")) %>%
  na.omit() %>%
  ungroup()

cbp_set <- c("#009E73", "#006299", "firebrick")

outcome_bars <- outcomes %>%
  mutate(disc_code = factor(disc_code, levels = arrange(., prop_unmatched)$disc_code)) %>%
  gather(-disc_code,-cluster, key = "Outcome", value = "prop") %>% 
  mutate(Outcome = factor(Outcome, 
                          levels = c("prop_matched_to_discipline", "prop_matched_to_alternative", "prop_unmatched"))) %>%
  ggplot(aes(x=disc_code, y=prop, fill = Outcome)) +
  geom_col() + 
  scale_fill_manual(labels = c("Matched to Discipline", "Matched to Alternative", "Unmatched"), values = cbp_set) + 
  ylab("Proportion of CMGs Ranking Discipline as First Choice") + 
  theme_pubr() + 
  facet_grid(.~cluster, scales="free_x", space = "free_x", labeller = as_labeller(cluster_labels)) + 
  theme(legend.position = "right", 
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 60, hjust = 1), 
        strip.text.x = element_text(size = 14)) 

outcome_bars
```

```{r}
cluster_palette <- c("#46B8DACC", "#EEA236BB", "#9632B8BB")

# Proportion unmatched in each cluster over time
cluster_unmatch_time <- discipline_trend_clust %>% 
  mutate(Cluster = str_replace(cluster, "Cluster ", "")) %>% 
  filter(!is.na(Cluster),
         Cluster != "NA") %>% 
  group_by(year, Cluster) %>% 
  dplyr::summarise(prop_unmatched = sum(unmatched) / sum(app_choice_1)) %>%
  ggplot(aes(x = year, y = prop_unmatched, color = Cluster)) +
  geom_line(size = 1.5) +
  scale_color_manual(values = cluster_palette) +
  geom_hline(yintercept = 0.25, linetype = 2, color = "grey") +
  geom_hline(yintercept = 0.20, linetype = 2, color = "grey") +
  geom_hline(yintercept = 0.15, linetype = 2, color = "grey") +
  geom_hline(yintercept = 0.10, linetype = 2, color = "grey") +
  geom_hline(yintercept = 0.05, linetype = 2, color = "grey") +
  scale_x_continuous(breaks=seq(2010, 2018,1)) +
  ylab("Proportion of Applicants Unmatched") +
  xlab("Year") + 
  theme_pubr() +
  theme(legend.title=element_text(size=13), 
          legend.text = element_text(size=12))# +
  #guides(color = guide_legend(override.aes = list(size=2)))

cluster_unmatch_time
```


```{r}
electives_clust_avg <- electives_clust %>% filter(year >= 2013) %>%
  group_by(cluster, disc_code) %>% 
  select(competitiveness, prop_unmatched, cluster, disc_code) %>% na.omit() %>%
  dplyr::summarise_all(., mean) %>% 
  ungroup() 

comp_unmatch_point <- electives_clust %>% 
  filter(year >= 2013) %>% 
  dplyr::rename("Cluster" = cluster) %>%
  ggplot(aes(x = competitiveness, y = prop_unmatched)) + 
  xlab("Competitiveness  (CMGs Ranking as Top Choice / Available Seats)") + ylab("Proportion of Applicants Unmatched") +
  geom_point(aes(color = Cluster), size = 0.5, alpha=0.3) +
  geom_text(data = electives_clust_avg, aes(label = disc_code, color = cluster), size = 3.5) +
  geom_smooth(aes(color = Cluster), method = "lm", se = FALSE, size = 0.25, alpha = 0.2) +
  stat_cor(size = 3.8, label.y.npc = 0.98) +
  stat_cor(aes(color = Cluster), size = 3.55, label.y.npc = 0.93) +
  scale_color_manual(values = cluster_palette) +
  theme_pubr() +
  theme(legend.title=element_text(size=13), 
          legend.text = element_text(size=12))
    

comp_unmatch_point
```

```{r, fig.width = 14, fig.height=10}
ggarrange(
  ggarrange(cluster_unmatch_time + theme(plot.margin = unit(c(1,3,2,1), "lines")), 
            comp_unmatch_point + theme(plot.margin = unit(c(1,1,2,1), "lines")), 
            ncol = 2, nrow = 1, labels = c("A", "B"), font.label = list(size = 24)),
  ggarrange(outcome_bars + theme(plot.margin = unit(c(1,0.5,1,2.5), "lines"), 
                                 legend.text=element_text(size=14), 
                                 legend.title = element_text(size=15)), 
            labels = c("C"), font.label = list(size = 24)),
  ncol = 1, nrow = 2) +
  ggsave("paper_figures/Figure2_Outcomes.png", width = 18, height = 13, dpi = 600)
```




###############################################################################################################################################

## Figure 3

Case study / illustrative example of derm vs plastics to highlight how differences in parallel planning influence match rates in equally competitive disciplines

```{r}

waffle_plot <-  function(data, clust,  years, as_prop = TRUE, nglyphs = 100, nrow = 5, glyph = "user", size = 6, colouring = FALSE, discipline_codes = FALSE, label_name = FALSE) {
  
  # function to round while preserving the sum 
  round_preserve_sum <- function(x) {
    y <- floor(x)
    indices <- tail(order(x-y), round(sum(x)) - sum(y))
    y[indices] <- y[indices] + 1
    y
  }

  if(discipline_codes == FALSE) {
    # gather and sum
    clust_sum <- data %>%
    filter(year %in% years,
           cluster == clust) %>%
    dplyr::summarise(matched_choice_1 = mean(matched_choice_1),
              matched_to_other = mean(matched_to_other),
              unmatched = mean(unmatched),
              prop_matched = matched_choice_1 / (matched_choice_1+matched_to_other+unmatched) * 100, 
              prop_matched_to_other = matched_to_other / (matched_choice_1+matched_to_other+unmatched) * 100,
              prop_unmatched = unmatched / (matched_choice_1+matched_to_other+unmatched) * 100) 
  } else { 
    # gather and sum
    clust_sum <- data %>%
    filter(year %in% years,
           disc_code %in% discipline_codes) %>%
    dplyr::summarise(matched_choice_1 = mean(matched_choice_1),
              matched_to_other = mean(matched_to_other),
              unmatched = mean(unmatched),
              prop_matched = matched_choice_1 / (matched_choice_1+matched_to_other+unmatched) * 100, 
              prop_matched_to_other = matched_to_other / (matched_choice_1+matched_to_other+unmatched) * 100,
              prop_unmatched = unmatched / (matched_choice_1+matched_to_other+unmatched) * 100) 
  }
  
  if(as_prop == TRUE){
    waffle_percent <- c("Matched to Discipline" = clust_sum$prop_matched, 
                    "Matched to Alternative" = clust_sum$prop_matched_to_other, 
                    "Unmatched" =  clust_sum$prop_unmatched)
    
    waffle_input <- round_preserve_sum(waffle_percent * nglyphs/100)
    
    print(waffle_percent)
    
  } else { 
    waffle_count <- c("Matched to Discipline" = clust_sum$matched_choice_1, 
                    "Matched to Alternative" = clust_sum$matched_to_other, 
                    "Unmatched" =  clust_sum$unmatched)
    
    waffle_input <- round_preserve_sum(waffle_count * nglyphs/100)
    
    print(waffle_count)
  }
  
  # generate waffleplot
  if(label_name != FALSE){
    
    if(is.logical(colouring)){
        waffplot <- waffle(waffle_input, rows = nrow, title = label_name, use_glyph = glyph, glyph_size = size)  
      } else {
        waffplot <- waffle(waffle_input, rows = nrow, title = label_name, use_glyph = glyph, glyph_size = size, colors = colouring)  
      }
    
  } else {
  
    if(discipline_codes == FALSE){
      if(is.logical(colouring)){
        waffplot <- waffle(waffle_input, rows = nrow, title = paste0("Cluster ", clust), use_glyph = glyph, glyph_size = size)  
      } else {
        waffplot <- waffle(waffle_input, rows = nrow, title = paste0("Cluster ", clust), use_glyph = glyph, glyph_size = size, colors = colouring)  
      }
    } else {
      if(is.logical(colouring)){
        waffplot <- waffle(waffle_input, rows = nrow, title = capture.output(cat(discipline_codes, sep = ", ")), use_glyph = glyph, glyph_size = size)  
      } else {
        waffplot <- waffle(waffle_input, rows = nrow, title = capture.output(cat(discipline_codes, sep = ", ")), use_glyph = glyph, glyph_size = size, colors = colouring)  
      }
    }
    
  }
  
  return(waffplot)
}
```



```{r}
years <- c(2013, 2013, 2015,2016,2017,2018,2019)

DP_waffle <- ggarrange(
  electives_clust %>%
    waffle_plot(., clust = FALSE, discipline_codes = "Derm", years, nglyphs=10, nrow = 1, size = 30, 
                glyph = c("user"), colouring = cbp_set, label_name = "Dermatology") + 
    theme(legend.text=element_text(size=16.5)),
  electives_clust %>%
    waffle_plot(., clust = FALSE, discipline_codes = "Plastics", years, nglyphs=10, nrow = 1, size = 30,
                glyph = c( "user"), colouring = cbp_set, label_name = "Plastic Surgery") +
    theme(legend.text=element_text(size=16.5)),
  nrow = 2, common.legend = TRUE, legend = "right"
) 
```



```{r, fig.width = 14, fig.height=10.5}
core_elective_labels <- c("mean_electives_within_discipline" = "Mean Electives Within Matched Discipline",
                          "log_unmatched_vs_alt" = "Unmatched / Matched to Alternative",
                          "electives_diversity" = "Diversity of Electives", 
                          "prop_backup" = "Frequency of Parallel Applications", 
                          "competitiveness" = "Competitiveness", 
                          "prop_unmatched" = "Proportion Unmatched",
                          "p_nomatch_other" = "Alternative Outcomes\n(Matched to Alternative vs Unmatched)")

my_comparisons <- list( c("Dermatology", "Plastic Surgery") )

structured_data_derm_plastics <- electives_clust %>%
  na.omit() %>%
  filter(year >= 2013,
         disc_code %in% c("Derm", "Plastics")) %>%
  mutate(Discipline = factor(disc_code, levels = c("Derm", "Plastics")),
         disc_class = ifelse(Discipline %in% c("Derm"), "Dermatology", "Plastic Surgery")) %>%
  select(year, Discipline, cluster, mean_electives_within_discipline,competitiveness, prop_unmatched,
         prop_backup, electives_diversity, disc_class, p_nomatch_other) %>%
  gather(key = "match_stat", value = "value", mean_electives_within_discipline, competitiveness, 
         prop_unmatched, prop_backup, electives_diversity, p_nomatch_other)


carms_boxplot_derm_plastics <- function(structured_data, metric){
  structured_data %>% 
    filter(match_stat == metric) %>%
    ggplot(aes(y = value, x = disc_class, fill = Discipline)) +
    geom_boxplot(outlier.color = 'grey', outlier.alpha = 0.8) + 
    scale_fill_manual(values = c("#EEA236BB", "#9632B8BB")) +
    facet_wrap(.~match_stat, scales = "free_y", labeller = as_labeller(core_elective_labels)) +
    xlab("") + ylab("") +
    stat_compare_means(comparisons = my_comparisons, size = 4, method = "t.test", label = "p.signif") +
    theme_pubr() +
    theme(legend.position = "none", 
          strip.text.x = element_text(size = 14)) 
}

ggarrange(
  ggarrange(DP_waffle + theme(plot.margin = unit(c(0,1,2,1.5), "lines")), 
            labels = "A", font.label = list(size = 26)),
  ggarrange(
    carms_boxplot_derm_plastics(structured_data_derm_plastics, "competitiveness") + theme(plot.margin = unit(c(1,1,0,1), "lines")),
    carms_boxplot_derm_plastics(structured_data_derm_plastics, "prop_unmatched") + theme(plot.margin = unit(c(1,1,0,1), "lines")),
    carms_boxplot_derm_plastics(structured_data_derm_plastics, "p_nomatch_other") + theme(plot.margin = unit(c(1,1,0,1), "lines")),
    ncol = 3, nrow = 1, labels = c("B", "C", "D"), font.label = list(size = 26)),
  ggarrange(
    carms_boxplot_derm_plastics(structured_data_derm_plastics, "mean_electives_within_discipline") + theme(plot.margin = unit(c(1,1,0,1), "lines")),
    carms_boxplot_derm_plastics(structured_data_derm_plastics, "electives_diversity") + theme(plot.margin = unit(c(1,1,0,1), "lines")),
    carms_boxplot_derm_plastics(structured_data_derm_plastics, "prop_backup") + theme(plot.margin = unit(c(1,1,0,1), "lines")),
    ncol = 3, nrow = 1, labels = c("E", "F", "G"), font.label = list(size = 26)), 
  ncol = 1, nrow = 3, heights = c(4,5.2,5)) +  
  ggsave("paper_figures/Figure3_Derm_Plastics.png", width = 18, height = 13.5, dpi = 600)
```



###############################################################################################################################################

## Figure 4

Characterizing unmatched risk and parellel planning across clusters 

```{r}
core_elective_labels <- c("mean_electives_within_discipline" = "Mean Electives Within Matched Discipline",
                          "electives_diversity" = "Diversity of Electives", 
                          "prop_backup" = "Frequency of Parallel Applications", 
                          "competitiveness" = "Competitiveness", 
                          "prop_unmatched" = "Proportion Unmatched",
                          "p_nomatch_other" = "Alternative Outcomes\n(Matched to Alternative vs Unmatched)")

my_comparisons <- list( c("A", "B"), c("B", "C"), c("A", "C") )


structured_data_clusters <- electives_clust %>%
  na.omit() %>%
  filter(year >= 2013) %>%
  select(year, cluster, mean_electives_within_discipline,competitiveness, prop_unmatched,
         prop_backup, electives_diversity, p_nomatch_other) %>%
  gather(key = "match_stat", value = "value", mean_electives_within_discipline, competitiveness, 
         prop_unmatched, prop_backup, electives_diversity, p_nomatch_other)

carms_boxplot_clusters <- function(structured_data, metric){
  structured_data %>% 
    filter(match_stat == metric) %>%
    ggplot(aes(y = value, x = cluster, fill = cluster)) +
    geom_boxplot(outlier.color = 'grey', outlier.alpha = 0.8) + 
    scale_fill_manual(values = cluster_palette) +
    facet_wrap(.~match_stat, scales = "free_y", labeller = as_labeller(core_elective_labels)) +
    xlab("") + ylab("") +
    stat_compare_means(comparisons = my_comparisons, size = 4, method = "t.test", label = "p.signif") +
    theme_pubr() +
    theme(legend.position = "none", 
          strip.text.x = element_text(size = 13.5)) 
}
```

```{r}
electives_clust_5yr <- electives_clust %>% filter(year >= 2013, cluster != "A") %>% na.omit() 

electives_clust_5yr_textsummary <- electives_clust_5yr %>%
  group_by(disc_code, cluster) %>%
  dplyr::summarise(competitiveness = mean(competitiveness),
                   electives_diversity = mean(electives_diversity),
                   prop_backup = mean(prop_backup),
                   mean_electives_within_discipline = mean(mean_electives_within_discipline),
                   p_nomatch_other = mean(p_nomatch_other))

p2 <- electives_clust_5yr %>% 
  na.omit() %>%
  ggplot(aes(x = electives_diversity, y = p_nomatch_other)) +
  xlab("Diversity of Electives") + ylab("Alternative Outcomes\n(Matched to Alternative vs Unmatched)") +
  geom_smooth(color = "light grey", method = "lm", se = FALSE, alpha = 0.6, size = 0.3) +
  geom_point(aes(color = cluster), shape = 16, alpha = 0.3, size = 0.7) +
  geom_text(data = electives_clust_5yr_textsummary, aes(x = electives_diversity, y = p_nomatch_other, color = cluster, label = disc_code), size = 3.5) +
  scale_color_manual(values = cluster_palette[2:3]) +
  stat_cor(size = 4, label.y.npc = 0.975) +
  theme_pubr()

p3 <- electives_clust_5yr %>% 
  na.omit() %>%
  ggplot(aes(x = prop_backup, y = p_nomatch_other)) +
  xlab("Frequency of Parallel Applications") + ylab("Alternative Outcomes\n(Matched to Alternative vs Unmatched)") +
  geom_smooth(color = "light grey", method = "lm", se = FALSE, alpha = 0.6, size = 0.3) +
  geom_point(aes(color = cluster), shape = 16, alpha = 0.3, size = 0.7) +
  geom_text(data = electives_clust_5yr_textsummary, aes(x = prop_backup, y = p_nomatch_other, color = cluster, label = disc_code), size = 3.5) +
  scale_color_manual(values = cluster_palette[2:3]) +
  stat_cor(size = 4, label.y.npc = 0.975) +
  theme_pubr()

```


```{r, fig.width = 14, fig.height=12}
ggarrange(
  ggarrange(
    carms_boxplot_clusters(structured_data_clusters, "competitiveness") + theme(plot.margin = unit(c(1,1,0,1), "lines")),
    carms_boxplot_clusters(structured_data_clusters, "prop_unmatched") + theme(plot.margin = unit(c(1,1,0,1), "lines")),
    carms_boxplot_clusters(structured_data_clusters, "p_nomatch_other") + theme(plot.margin = unit(c(1,1,0,1), "lines")),
    ncol = 3, nrow = 1, labels = c("A", "B", "C"), font.label = list(size = 26)),
  ggarrange(
    carms_boxplot_clusters(structured_data_clusters, "mean_electives_within_discipline") + theme(plot.margin = unit(c(1,1,0,1), "lines")),
    carms_boxplot_clusters(structured_data_clusters, "electives_diversity") + theme(plot.margin = unit(c(1,1,0,1), "lines")),
    carms_boxplot_clusters(structured_data_clusters, "prop_backup") + theme(plot.margin = unit(c(1,1,0,1), "lines")),
    ncol = 3, nrow = 1, labels = c("D", "E", "F"), font.label = list(size = 26)), 
  ggarrange(
    p2 + theme(plot.margin = unit(c(1,4,0.5,2), "lines")),
    p3 + theme(plot.margin = unit(c(1,4,0.5,2), "lines")),
    common.legend = TRUE, legend = "none", nrow = 1, ncol = 2, labels = c("G", "H"), font.label = list(size = 26), widths = c(1,1)),
  ncol = 1, nrow = 3, heights = c(5, 5, 6.4)) +
  ggsave("paper_figures/Figure4_Cluster_Parallel.png", width = 18, height = 15.5, dpi = 600)

```



# Figure 5
Looking at program selection behaviours with regards to on-site electives


```{r}
core_elective_labels <- c("prop_ranked_program_elective" = "Proportion Ranked with Program Elective",
                          "prop_matched_program_elective" = "Proportion Matched with Program Elective")

structured_data_program <- electives_clust %>%
  na.omit() %>%
  filter(year >= 2013) %>%
  select(year, cluster, prop_ranked_program_elective, prop_matched_program_elective) %>%
  gather(key = "match_stat", value = "value", prop_ranked_program_elective, prop_matched_program_elective) %>%
  mutate(match_stat = factor(match_stat, levels = c("prop_ranked_program_elective", "prop_matched_program_elective")))


### Correlation Plots
electives_clust_5yr <- electives_clust %>% filter(year >= 2013) %>% na.omit() 

electives_clust_5yr_textsummary <- electives_clust_5yr %>%
  group_by(disc_code, cluster) %>%
  dplyr::summarise(competitiveness = mean(competitiveness),
                   electives_diversity = mean(electives_diversity),
                   prop_backup = mean(prop_backup),
                   prop_ranked_program_elective = mean(prop_ranked_program_elective),
                   prop_matched_program_elective = mean(prop_matched_program_elective),
                   mean_electives_within_discipline = mean(mean_electives_within_discipline),
                   p_nomatch_other = mean(p_nomatch_other))

ranked_electives <- electives_clust_5yr %>% 
  na.omit() %>%
  ggplot(aes(x = prop_ranked_program_elective, y = mean_electives_within_discipline)) +
  xlab("Proportion Ranked with Program Elective") + ylab("Mean Electives within Discipline") +
  geom_smooth(color = "light grey", method = "lm", se = FALSE, alpha = 0.6, size = 0.3) +
  geom_point(aes(color = cluster), shape = 16, alpha = 0.3, size = 0.7) +
  geom_text(data = electives_clust_5yr_textsummary, aes(x = prop_ranked_program_elective, y = mean_electives_within_discipline, 
                                                        color = cluster, label = disc_code), size = 3.2) +
  scale_color_manual(values = cluster_palette[1:3]) +
  stat_cor(size = 4, label.y.npc = 0.975) +
  theme_pubr()

matched_electives <- electives_clust_5yr %>% 
  na.omit() %>%
  ggplot(aes(x = prop_matched_program_elective, y = mean_electives_within_discipline)) +
  xlab("Proportion Matched with Program Elective") + ylab("Mean Electives within Discipline") +
  geom_smooth(color = "light grey", method = "lm", se = FALSE, alpha = 0.6, size = 0.3) +
  geom_point(aes(color = cluster), shape = 16, alpha = 0.3, size = 0.7) +
  geom_text(data = electives_clust_5yr_textsummary, aes(x = prop_matched_program_elective, y = mean_electives_within_discipline, 
                                                        color = cluster, label = disc_code), size = 3.2) +
  scale_color_manual(values = cluster_palette[1:3]) +
  stat_cor(size = 4, label.y.npc = 0.975) +
  theme_pubr()

ranked_diversity <- electives_clust_5yr %>% 
  na.omit() %>%
  ggplot(aes(x = prop_ranked_program_elective, y = electives_diversity)) +
  xlab("Proportion Ranked with Program Elective") + ylab("Diversity of Electives") +
  geom_smooth(color = "light grey", method = "lm", se = FALSE, alpha = 0.6, size = 0.3) +
  geom_point(aes(color = cluster), shape = 16, alpha = 0.3, size = 0.7) +
  geom_text(data = electives_clust_5yr_textsummary, aes(x = prop_ranked_program_elective, y = electives_diversity, color = cluster, label = disc_code), size = 3.2) +
  scale_color_manual(values = cluster_palette[1:3]) +
  stat_cor(size = 4, label.y.npc = 0.975) +
  theme_pubr()

matched_diversity <- electives_clust_5yr %>% 
  na.omit() %>%
  ggplot(aes(x = prop_matched_program_elective, y = electives_diversity)) +
  xlab("Proportion Matched with Program Elective") + ylab("Diversity of Electives") +
  geom_smooth(color = "light grey", method = "lm", se = FALSE, alpha = 0.6, size = 0.3) +
  geom_point(aes(color = cluster), shape = 16, alpha = 0.3, size = 0.7) +
  geom_text(data = electives_clust_5yr_textsummary, aes(x = prop_matched_program_elective, y = electives_diversity, color = cluster, label = disc_code), size = 3.2) +
  scale_color_manual(values = cluster_palette[1:3]) +
  stat_cor(size = 4, label.y.npc = 0.975) + ylim(0.55, 1.05) +
  theme_pubr()
```

```{r, fig.width = 14, fig.height=8}
ggarrange(
  ggarrange(
    carms_boxplot_clusters(structured_data_program, "prop_ranked_program_elective") + 
              theme(plot.margin = unit(c(1, 1.5, 1, 0.5), "lines")), 
    carms_boxplot_clusters(structured_data_program, "prop_matched_program_elective") + 
              theme(plot.margin = unit(c(0.5, 1.5, 1, 0.5), "lines")), 
    labels = c("A", "B"), font.label = list(size = 24), nrow = 2, ncol = 1),
  ggarrange(
    ggarrange(ranked_electives + theme(plot.margin = unit(c(1,1,1,1), "lines")), 
              ranked_diversity + theme(plot.margin = unit(c(1,1,1,1), "lines")), 
              ncol = 2, widths = c(1,1), labels = c("C", "E"), font.label = list(size = 24), common.legend = TRUE, legend = "none"),
    ggarrange(matched_electives + theme(plot.margin = unit(c(1,1,1,1), "lines")), 
              matched_diversity + theme(plot.margin = unit(c(1,1,1,1), "lines")), 
              ncol = 2, widths = c(1,1), labels = c("D", "F"), font.label = list(size = 24), common.legend = TRUE, legend = "none"),
    nrow = 2, ncol = 1, heights = c(1,1)),
  nrow = 1, ncol = 2, widths = c(1,2.4)) + 
  ggsave("paper_figures/Figure_5_ranked_electives.png", width = 16, height = 9, dpi = 600)

```

