---
title: "4 - Supplemental Analysis - Figures S1 and S3"
output:
  html_document:
    df_print: paged
---

Supplemental analysis that did not make it into the primary figures. 

```{r}
library(MASS)
library(tidyverse)
library(readr)
library(pheatmap)
library(ggpubr)
library(ggplotify)
library(betareg)
library(RColorBrewer)
library(statmod)
```

```{r}
electives <- read.csv("electives.csv", stringsAsFactors = FALSE)

electives_clust <- tibble("A" = c("Microbio", "Gen Path", "Nuclear", "Pub Health","Neuro Path","Peds Neuro",
                          "Heme Path", "Anat Path","Internal","Family", "Med Gen", NA),
                  "B" = c("Psych", "Rad Onc", "OBGYN", "Peds", "Neuro", "Physiatry", "Anesth",
                          "Rads", "Gen Surg", "Vasc Surg", "Derm", "Emerg"),
                  "C" = c("Cardiac Surg", "Ophthal",  "Neuro Surg", "Urology",
                          "Ortho", "ENT", "Plastics", NA, NA, NA, NA, NA)) %>%
 gather(key = "cluster", value = "disc_code") %>% na.omit() %>%
 right_join(.,electives, by="disc_code")
```


### Figure S1 - Cluster Features

```{r, fig.height=7, fig.width=10}
cluster_palette <- c("#46B8DACC", "#EEA236BB", "#9632B8BB")
tsne_data <- read_csv("tsne_electives.csv")

electives_mat <- tsne_data %>% 
  dplyr::select(-c("discipline", "discipline_full", "disc_code", "prop_unmatched", "TSNE1", "TSNE2")) %>% data.matrix()
rownames(electives_mat) <- tsne_data %>% pull(disc_code)

# get dendrogram
corelectives <- electives_mat %>% t() %>% cor()
electives_dist <- as.dist((1-corelectives)/2)
electives_hc <- hclust(electives_dist, method = "complete")

```


```{r, fig.height=7, fig.width=10}
# matrix for heatmap
electives_mat_with_unmatched <- tsne_data %>%
  dplyr::select(-c(disc_code, discipline, discipline_full, TSNE1, TSNE2, electives_clusters)) %>% data.matrix
rownames(electives_mat_with_unmatched) <- tsne_data$disc_code

col_labels <- c("Prop Matched with \nDiscipline Elective", "Prop Matched with \nProgram Elective", 
                "Prop Ranked with \nDiscipline Elective", "Prop Ranked with \nProgram Elective",
                "Mean Electives within\nMatched Discipline", "Mean Electives outside\nMatched Discipline",
                "Mean Other Disciplines with\n Completed Electives", "Prop Matched with >=3\n Discipline Electives",
                "Competitiveness", "Frequency of\nParallel Applications", "Prop Applicants Unmatched")

# Plot the heatmap
electives_heatmap <- pheatmap(electives_mat_with_unmatched, color = colorRampPalette(rev(brewer.pal(n = 11, name = "RdYlBu")))(100), 
                              scale = "column", angle_col = "45", cutree_rows = 3, cutree_cols = 4, fontsize = 12, fontsize_row = 10, 
                              fontsize_col = 9, fontsize_number = 10, legend = TRUE, legend_breaks = c(-3,-2,-1,0,1,2,3,3), 
                              legend_labels = c("-3","-2","-1"," 0"," 1"," 2"," 3","Z-score\n\n\n"), 
                              cluster_rows = electives_hc, labels_col = col_labels, silent = TRUE) 

ggarrange(as.grob(electives_heatmap)) + 
  ggsave("paper_figures/Figure_S1_cluster_heatmap.png", width = 12, height = 8, dpi = "retina")
```



## Multivariate Regression 

Running multiple linear regression and beta regression to see which terms are still significant.
Also running stepwise BIC with all terms and all interactors to look for significant interactions influencing unmatched rates

```{r}
electives <- read_csv("electives.csv")

# match input from electives data
match_input <- electives %>% 
  dplyr::select(-c(year, discipline, discipline_full, disc_code)) %>% 
  na.omit()

```

```{r}
lin <- lm(prop_unmatched ~ ., data=match_input)

summary(lin)
```

```{r}
beta_input <- match_input %>% mutate(prop_unmatched = (prop_unmatched*(dim(match_input)[1] - 1) + 0.5) / dim(match_input)[1])
beta <- betareg(prop_unmatched ~ ., data = beta_input)

summary(beta)
```


```{r, results='hide',fig.keep='all'}
### Stepwise BIC approach 
stepwise_match_input <- lm(prop_unmatched ~ 
  (prop_backup + competitiveness + prop_ranked_program_elective + prop_matched_program_elective + mean_other_disciplines + 
     prop_ranked_discipline_elective + prop_matched_discipline_elective + 
     mean_electives_outside_discipline + mean_electives_within_discipline + prop_matched_3_discipline_elective)^2, data = match_input)

stepwise_match_bic <- stepAIC(stepwise_match_input, k = log(nrow(match_input)))


# compare AICs 
model <- c("match_lm", "match_lm_stepwise_BIC", "match_beta")
AIC <- c(AIC(lin), AIC(stepwise_match_bic), AIC(beta))
BIC <- c(BIC(lin), BIC(stepwise_match_bic), BIC(beta))
scores <- data.frame(model, AIC, BIC); scores
```

```{r}
summary(stepwise_match_bic)
```

After running stepwise multiple regression the significant interactor terms are: 
  competitiveness : mean_electives_within_discipline
  competitiveness : prop_backup
We will visualize these interactions in figure S3


## Figure S3

Interaction plots of: 
    - Electives within Discipline vs Unmatched    [ interactor: Competitiveness ]
    - Frequency of Parallel Applications vs Unmatched          [ interactor: Competitiveness ]
   

```{r, fig.height=5, fig.width=14}

electives_clust_since2013 <- electives_clust %>% filter(year >= 2013) %>% na.omit()

comp_quantiles <- electives_clust_since2013 %>%
  pull(competitiveness) %>% 
  quantile(., probs = c(0.33, 0.66), na.rm = TRUE)

# electives over competitiveness levels
p1 <- electives_clust_since2013 %>%
  mutate(Competitiveness = if_else(competitiveness < comp_quantiles[[1]], "Low", 
                                 if_else(competitiveness < comp_quantiles[[2]], "Med", "High"))) %>%
  mutate(Competitiveness = factor(Competitiveness, levels = c("High", "Med", "Low"))) %>% 
  na.omit() %>%
  ggplot(aes(x = mean_electives_within_discipline, y = prop_unmatched)) +
  xlab("# of Electives in Matched Discipline") + ylab("Proportion of Applicants Unmatched") +
  geom_point(aes(color = Competitiveness), size = 1.2) +
  geom_smooth(aes(color = Competitiveness), method = "lm", se = FALSE) +
  scale_color_brewer(type = "qual", palette = 6, name = "Competitiveness") +
  stat_cor(size = 3.2, label.y.npc = 0.975) +
  stat_cor(aes(color = Competitiveness), size = 3, label.y.npc = 0.925) +
  theme_pubr()


# backup over competitiveness levels
p2 <- electives_clust_since2013 %>%
  mutate(Competitiveness = if_else(competitiveness < comp_quantiles[[1]], "Low", 
                                 if_else(competitiveness < comp_quantiles[[2]], "Med", "High"))) %>%
  mutate(Competitiveness = factor(Competitiveness, levels = c("High", "Med", "Low"))) %>% 
  na.omit() %>%
  ggplot(aes(x = prop_backup, y = prop_unmatched)) +
  xlab("Frequency of Parallel Applications") + ylab("Proportion of Applicants Unmatched") +
  geom_point(aes(color = Competitiveness), size = 1.2) +
  geom_smooth(aes(color = Competitiveness), method = "lm", se = FALSE) +
  scale_color_brewer(type = "qual", palette = 6, name = "Competitiveness") +
  stat_cor(size = 3.2, label.y.npc = 0.975) +
  stat_cor(aes(color = Competitiveness), size = 3, label.y.npc = 0.925) +
  theme_pubr()


ggarrange(
  p1, p2, nrow = 1, ncol = 2, 
  labels = c("A", "B"), font.label = list(size = 22)) +
  ggsave("paper_figures/Figure_S3_interaction_plots.png", width = 16, height = 5, dpi = "retina")

```


